I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-23 Paper/proofs.tex
fix typos and others

\begin{lemma}[Sign]
	\label{lemma1}
	Let $L,L'$  consecutive layers of a DNN without compensation, 
	and $m\in L$ $n\in L'$. Then if both 
	$W_{nm} \neq 0$ and $S(n) \neq 0$, then 
	$S(m)=S(n)\mathrm{Sign}(W_{n m})$.
\end{lemma}

@@

	Hence there is a non zero path from $m$ to $z$: $(m n) \pi$, which is of sign 
	$S(n)\mathrm{Sign}(W_{n m})$. As there is no compensation, $S(m)=S(n)\mathrm{Sign}(W_{n m})$.
	\hfill $\square$
\end{proof}

For a node $n$, we use $n_s$ to denote $S(n)\cdot n$. 
Notice that for $S(n)=1$, $n_s$ gets maximal value whenever $n$ gets maximal value; 
while for $S(n)=-1$, $n_s$ gets maximal value whenever $n$ gets minimal value (and vice versa). For $S(n)=0$, $n_s=0$ and thus always reach his minimal and maximal.

@@

%	\begin{lemma}
	%		\label{lemma2}
	%		Let $L,L'$ be consecutive layers of a compensation free DNN, and $n \in L'$. 
	%		Then:
	%		$$ S(n) \max(S(n) n)=\sum_{m \in L}W_{m n} S(m) \max(S(m) \hat{m})+bias_n \text{ and }$$
	%		$$S(n) \max(S(n)\hat{n}_s)=\ReLU(\sum_{m \in L}W_{m n} S(n)\max(S(n)\hat{m}_s)+bias_n)$$
	%		
	%		Similarly for minimal value.
	%		
	%		%	for minimal value,	
	%		%	$ \min(n)=\sum_{m \in L}W_{m n)\min(\)+bias_n $ and
		%			%	$ \min(\hat{n}_s)=ReLU(\sum_{m \in L}W_{m n)\min(\hat{m}s}+bias_n)$
		%		\end{lemma}
	%		
	%		\begin{proof}
		%			We choose $\boldsymbol{x}^*$ maximizing each $a_s \in \{a,-a\}$ in the input layer $L=L_0$,
		%			that is setting the value for node $a$ be $\max(a)$ for $S(a)=1$ and $\min(a)$ for $S(a)=-1$.
		%			At configuration $\boldsymbol{x}^*$, every $a_s(\boldsymbol{x}^*))=\max_{\boldsymbol{x}} {a_s(\boldsymbol{x}))}=\max(a_s)$ is maximized.  
		%			
		%			Consider any $b$ in the first hidden layer $L_1$.
		%			Assume first that $S(b)=+1$.
		%			We have $\Val(b_s)= \Val(b) = \sum_{a \in L_0} W_{ab} \Val(a) + bias_b$.
		%			If every $W_{ab} \Val(a)$ is maximized in the same time, 
		%			then $\Val(b_s)= \Val(b)$ is also maximized. 
		%			
		%			We claim that this is the case at 
		%			configuration $\boldsymbol{x}^*$.
		%			Indeed, if $W_{ab}=0$, any $\Val(a)$ can be chosen to maximize 
		%			$W_{ab}\Val(a)$. If $W_{ab}>0$ then $S(a) = +1$ by Lemma \ref{lemma1}, since
		%			$S(b)=+1$. We have $\max(a_s) = \max(a)$, maximizing $W_{ab}\Val(a)$.
		%			Finally, if $W_{a b}<0$ then $S(a_s) = -1$, and we have 
		%			$\max(a_s) = min(a)$, maximizing $W_{ab}\Val(a)$.
		%			
		%			Hence $b$ reaches is maximum in configuration $\boldsymbol{x}^*$, and the value satisfies: 
		%			$$\max(b)=\sum_{m \in L}W_{a b} \max(a_s)+bias_b$$
		%			
		%			The case $S(b)=-1$ is symmetric:
		%			$\max(b_s)= -\min(b)$ is reached when minimizing $b$, which is also 
		%			satisfied at $\boldsymbol{x}^*$.
		%			In both case, $\Val(b_s)$ is maximized at $\boldsymbol{x}^*$. Notice that $\boldsymbol{x}^*$ does not depend upon $b$, so choosing $\boldsymbol{x}^*$ uniformly maximize all $\Val(b_s)$.
		%			
		%			This implies that $\ReLU(b_s)$ is maximized at $\boldsymbol{x}^*$, and its value is 
		%			$\ReLU(\sum_{m \in L}W_{a }\max(a_s)+bias_b)$, again uniformly reached over all $b$ at configuration $\boldsymbol{x}^*$.
		%			
		%			We proceed by induction over every layer of the DNN till the output layer, proving that 
		%			$\max(z_s)=\max(z)$ is reached at configuration $\boldsymbol{x}^*$.
		%			
		%			The case for $\min$ is similar, with another input vector $\boldsymbol{x}^\sharp$ instead of $\boldsymbol{x}^*$.
		%			\hfill $\square$
		%		\end{proof}

@@

	Notice that for all $x$, for both $S(c)=1$ or $S(c)=-1$, these two bounds are found in Lemma \ref{lemma2},
	from configuration $\boldsymbol{x}^*$ and $\boldsymbol{x}^\sharp$.

@@

	%\section{Theorem 2}
	%
	%We now show that if all intermediates nodes that are on compensating pairs are opened as MILP nodes, then MILP will be correct. First, we do the proof with 3 layers, assuming (without loss of generality) a unique output node $z$.
	%
	%\begin{theorem}
	%	\label{no_diamond_2}
	%	Using MILP method, if every nodes $b_j$ in any compensating path pair
	%	 $(\pi,\pi')$ is encoded as a binary/integer variable, then the upper and lower 
	%	 bounds computed by MILP are the exact max and min value of $z$.
	%\end{theorem}
	%
	%By symetry, we only show the max side. 
	%
	%
	%
	%\begin{definition}
	%	Let $K$ be the set of all input nodes $a_k$. 
	%	We define a decomposition $K=I\sqcup J$ as follows:  
	%	\begin{itemize}
		% \item $k \in I$  if
		% \begin{enumerate}
			%	 \item every path from $a_k$ to $z$ has weight $\geq 0$, or
			%	\item every path from $a_k$ to $z$ has weight $\leq 0$.
			% \end{enumerate}
		%That is, $a_k$ is not a source of a compensating pair.
		%	\item $k \in J$ if $k \notin I$, that is there exists two paths $\pi,\pi'$ from $a_k$, 
		%	one with positive and one with negative weight.
		%\end{itemize}
		%\end{definition} 
		%
		%
		%\begin{lemma} \label{lem:open_node}
		%	A node $b$ in the hidden layer will not be on a compensating pair iff one of the following two happens:
		%	\begin{enumerate}
			%	 \item $W_{b,z}=0$, or
			%	 \item For every input node $j\in J$, we have $W_{a_j,b}=0$.
			%	\end{enumerate}
		%	We denote $B_{pure}$ the set of such nodes $b$ such that at least one of the above holds.
		%\end{lemma}
		%
		%\begin{proof}
		%	First, we show that if either one of 1,2 happens, then $b_i$ will not be opened. If 1), it is obvious. For 2, we reason by contradiction: assume there is a pair of compensating paths 	$(\pi,\pi')$ starting with $a$, with $k$ in $\pi$ and weight$(\pi) > 0$. It means that $a \in J$. A contradiction as 2) $W_{a_j,b}=0$ implies weight$(\pi)=0$.
		%	
		%	Second, we show that if neither 1 nor 2 hold, then $b$ will be on a compensating path.
		%	Because 2) does not hold, there is a $j \in J$ with $W_{a_j,b} \neq 0$, say $>0$.
		%	Because 1) does not hold, $W_{b,z} \neq 0$, say $>0$.
		%	Now, by definition of $J$, there is a pair of compensating paths $\pi,\pi'$ 
		%	from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
		%	Then $((a_j,b,z), \pi')$ is also a compensating pair.
		%\end{proof}
		%
		%Consider the sign of nodes function from Definition \ref{sign_of_nodes}. Because now we allow compensating paths, we cannot define this function over all nodes.
		%
		%
		%\begin{definition}\label{sign_of_nodes_in_I}
		%	We define a partial sign function $S$ over nodes $n$ such that : 	
		%	\begin{enumerate} 
			%		 \item all paths from $n$ to $z$ have 0 weight, and then $S(n)=0$; 
			%		 \item all paths from $n$ to $z$ have non-negative weight, and at least one path has a positive weight, and then $S(n)=1$; 
			%		 \item all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight, and then $S(n)=-1$.
			%		 \item Otherwise $S(n)$ is undefined.
			%	\end{enumerate}
		%\end{definition}	
		%	
		%Notice that $S$ is defined on all nodes $a_i$ with $i \in I$ (it can also be defined trivially for all nodes $b$ as there is a unique path to $z$, and at $z$ also). Hence it is undefined only for nodes $a_j$ with $j \in J$.
		%
		%
		%We denote $a_S$ for any subset $S\subseteq K$ to refer the input vector $\langle a_k\rangle_{a_k\in S}$. We also denote $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}_K$ and $z=z(\boldsymbol{x}_K)=z(\boldsymbol{x}_I,\boldsymbol{x}_J)$.
		%Consider $\boldsymbol{x}_I^*$ the input vector such that for all $i \in I$, the value for $a_i$ gets its maximal value (if $S(a_i)=1$) or $a_i$ gets its minimal value (if $S(a_i)=-1$).
		%
		%\begin{lemma} \label{lem:reach_max}
		%	$max_a (weight_{a}(z)) = max_{\{a \mid \boldsymbol{x}_I=\boldsymbol{x}^*_I\}} (weight_{a}(z))$
		%	
		%	Further, for every intermediate node $b$ in the hidden layer, for any valuation $\boldsymbol{x}^0_J$, 
		%	we have $max_{\{a \mid \boldsymbol{x}_J=\boldsymbol{x}^0_J\}} (weight_{a}(b)) = max_{\{a \mid \boldsymbol{x}_J=\boldsymbol{x}^0_J,\boldsymbol{x}_I=\boldsymbol{x}_I^*\}} (weight_{a}(b))$.	
		%\end{lemma}
		%
		%\begin{proof}
		%	For the first statement, we use 
		%	$$max_{\boldsymbol{x}_K} (weight_{\boldsymbol{x}_K}(z)) = max_{\boldsymbol{x}_J} max_{\boldsymbol{x}_I} (weight_{\boldsymbol{x}_I,\boldsymbol{x}_J}(z))$$
		%	
		%	Now, for any fixed input $\boldsymbol{x}^0_J$, we can regard $weight_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $\boldsymbol{x}^0_J$ %and their propagation in hidden layers 
		%	as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Theorem \ref{th1} (make a more precise corollary) to get that $z(\cdot,\boldsymbol{x}^0_J)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and we are done.
		%	
		%	The second statement is simpler because for each fixed partial input $\boldsymbol{x}^0_J$, 
		%	$$b= B^0 +\sum_{a_i\in I} W_{a_i, b} a_i,$$ where $B^0$ is a constant which is the sum of the term including bias and $\boldsymbol{x}^0_J$. Applying Lemma \ref{lemma1}, we obtain the statement.
		%\end{proof}
		%
		%\subsection{MILP abstraction}
		%
		%Consider now an MILP abstraction using the abstraction from \cite{MILP}, 
		%where each variable $\alpha_b$ for the ReLU from $b^-$ to $b^+$ is:
		%\begin{itemize}
		%	\item linear  (i.e. $\alpha_b \in [0,1]$) for $b \in B_{pure}$
		%	\item binary/integer (i.e. $\alpha_b \in \{0,1\}$) for $b \notin B_{pure}$
		%\end{itemize}
		%
		%
		%\begin{lemma}\label{lem:AppB=ReLU}
		%	Let $\mathrm{AppB}$ be the upper bound of lower bound approximation functions of LP for $\ReLU$ function with known upper and lower bound (numbers) $u,l$. Then when $x$ reaches the extremal values (maximum or minimum values, $u$ or $l$), then $\mathrm{AppB}(x)=\ReLU(x)$.
		%	
		%\end{lemma}
		%
		%\begin{proof}
		%	This is simply by definition of LP approximation functions.
		%\end{proof}
		%
		%Notice that there is only one layer of ReLUs, in the hidden layer.
		%Denote by $\mathrm{UB}$ the maximal bound on $z$ considering the MILP constraints above, with ReLUs being either linear or binary depending on whether $b \in B_{pure}$ or not.
		%
		%We want to show that $\mathrm{UB} = \max z$. As the MILP abstraction is a sound overapproximation, 
		%it suffices to show that $\mathrm{UB}\leq \max c$.
		%
		%For any $b_h\in B_{pure}$, we have $b_h=\sum_{a_i\in I}W_{b_ha_i}+B_h$ ($B_h$ is the bias). 
		%Denote by $B_{open}$ the set of nodes $b$ for which $\alpha_b$ is binary, i.e. 
		%$b \notin B_{pure}$. %For any subdomain $D \subseteq B_{open}$, 
		%Let $\boldsymbol{x}^0_J$ and $\boldsymbol{x}^0_I$ be fixed inputs, and consider the associated upper bound 
		%$\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I}$. We have:
		%%We show that this upper bound does not exceed the maximal value of $z$: 
		%
		%\begin{align*}
		%	\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I} = B_z + \sum_{b\in B_{open}} W_{b,z}\ReLU(weight_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)) + \sum_{b\in B_{pure}} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)).
		%\end{align*} where $\mathrm{AppB}(weight_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b))$ is the upper bound of 
		%the (LP) approximation of $weight_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)$ if $W_{b,z}>0$, and the lower bound
		%of its (LP) approximation if $W_{b,z}<0$. 
		%
		%By Lemma \ref{lem:reach_max}, for any fixed input $\boldsymbol{x}^0_J$, for all nodes $b$ in the hidden layer, if $S(b)=\mathrm{sign}(W_{zb})=1$, then 
		%$weight_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(b)$ will get its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$,
		%and if $S(b)=\mathrm{sign}(W_{zb})=-1$, then $weight_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}(b)$ will get its minimal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$.
		%Notice that $\mathrm{ReLU}$ and $\mathrm{AppB}$ are both non decreasing functions, and thus:
		%
		%\begin{align*}
		%	\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I} \leq B_z + \sum_{b\in B_{open}} W_{b,z}\ReLU(weight_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)) +
		%	\sum_{b\in B_{pure}} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)).
		%\end{align*} 
		%
		%Added FIX:
		%Consider $B_{pure}=B^1 \cup B^2$, with $b \in B^1$ iff
		%$W_{b,z} = 0$ and $b \in B^2$ iff $W_{b,z} \neq 0$.
		%Hence:
		%
		%$$\sum_{b\in B_{pure}} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)) = 
		%\sum_{b\in B^2} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b))$$
		%
		%For $b \in B^2$, we have $W_{a_j,b}=0$ for all $j \in J$
		%by definition of $B_{pure}$ (Lemma 4.), hence
		%
		%$$\sum_{b\in B^2} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(b)) = 
		%\sum_{b\in B^2} W_{b,z} \mathrm{AppB}(weight_{\boldsymbol{x}^*_I}(b))
		%=\sum_{b\in B^2} W_{b,z} \mathrm{ReLU}(weight_{\boldsymbol{x}^*_I}(b))
		%$$ as $\mathrm{AppB}(weight_{\boldsymbol{x}^*_I}(b))=\mathrm{ReLU}(weight_{\boldsymbol{x}^*_I}(b))$ since 
		%$weight_{\boldsymbol{x}^*_I}(b)$ is an extremal value (minimum or maximal) and 
		%$\mathrm{AppB}$ equals $\mathrm{ReLU}$ for extremal values.
		%
		%Hence we have: \begin{align*}
			%	\mathrm{UB}_{\boldsymbol{x}^0_J,\boldsymbol{x}^0_I}  \leq & B_z+\sum_{b}W_{zb}\ReLU(b(\boldsymbol{x}_I^*,\bar{a}_J))
			%	= \max_{\boldsymbol{x}_I} z(\boldsymbol{x}_I,\bar{a}_J) \\
			%	\leq & \max_{\boldsymbol{x}_J}\max_{\boldsymbol{x}_I} z(\boldsymbol{x}_I,\boldsymbol{x}_J) 	= \max_{\boldsymbol{x}_K} weight_{\boldsymbol{x}_K}(z)
			%\end{align*}
			%
			%Considering the maximal over all inputs $\boldsymbol{x}_J,\boldsymbol{x}_I$, we obtain 
			%
			%$$\mathrm{UB} = max_{\boldsymbol{x}_J,\boldsymbol{x}_I} \mathrm{UB}_{\boldsymbol{x}_J,\boldsymbol{x}_I} \leq \max_{\boldsymbol{x}_J,\boldsymbol{x}_I} weight_{\boldsymbol{x}_J,\boldsymbol{x}_I}(z)$$ And this us what we want to show.

>>>>

\begin{lemma}[Sign]
	\label{lemma1}
	Let $L,L'$ be consecutive layers of a DNN without compensation, 
	and $m\in L$ $n\in L'$. If
	$W_{nm} \neq 0$ and $S(n) \neq 0$, then 
	$S(m)=S(n)\mathrm{Sign}(W_{n m})$.
\end{lemma}

@@

	So there is a non zero path from $m$ to $z$: $(m n) \pi$, whose sign is 
	$S(n)\mathrm{Sign}(W_{n m})$. As there is no compensation, $S(m)=S(n)\mathrm{Sign}(W_{n m})$.
	\hfill $\square$
\end{proof}

%For a node $n$, we use $n_s$ to denote $S(n)\cdot n$. 
%Notice that for $S(n)=1$, $n_s$ gets maximal value whenever $n$ gets maximal value; 
%while for $S(n)=-1$, $n_s$ gets maximal value whenever $n$ gets minimal value (and vice versa). For $S(n)=0$, $n_s=0$ and thus always reach his minimal and maximal.

@@

@@

	Notice that for all $x$, for either $S(c)=1$ or $S(c)=-1$, these two bounds are found in Lemma \ref{lemma2},
	from configuration $\boldsymbol{x}^*$ and $\boldsymbol{x}^\sharp$.

@@
